Highly Available Transactions: Virtues and Limitations\\(Extended Version)

### 2013-07-25 writeup/hat-analysis.tex
next -> later

Atomicity, informally guaranteeing that either all or none of
transactions' effects should succeed, is core to ACID
guarantees. Although, at least by the ACID acronym, atomicity is not an
``isolation'' property, atomicity properties also restrict the updates
visible to other transactions. Accordingly, here, we consider the
\textit{isolation} effects of atomicity, which we call \textbf{Read
  Atomic (RA)} isolation.  Under RA, once some of the effects of a
transaction $T_i$ are observed by another transaction $T_j$,
thereafter, all effects of $T_i$ are observed by $T_j$. That is, if a
transaction $T_j$ reads a version of an object that transaction $T_i$
wrote, then a later read by $T_j$ cannot return value whose next
version is installed by $T_i$. Together with item cut isolation, RA
prevents Read Skew anomalies (Berenson et al.'s A5A) and is useful in
several contexts such as maintaining foreign key constraints,
consistent global secondary indexing, and maintenance of derived
data. In the example below, under RA, because $T_2$ has read $T_1$'s
write to $y$, $T_2$ must observe $b=c=1$ (or later versions for each
key):
\begin{align*}
\small
T_1 &: w_x(1)~w_y(1)~w_z(1)
\\T_2 &: r_x(a)~r_y(1)~r_x(b)~r_z(c)~\\[-1.5em]
\end{align*}
$T_2$ can also observe $a=\bot$, $a=1$, or a later version of $x$. In
the hierarchy of existing isolation properties, we place RA below
Adya's \textit{PL-2L} (as it does not necessarily enforce transitive
read-write dependencies) but above Read Committed ($PL-2$). Notably,
RA requires disallows reading intermediate writes (Adya's $G1b$):
observing all effects of a transaction implicitly requires observing
the final (committed) effects of the transaction as well.

>>>>

Atomicity, informally guaranteeing that either all or none of
transactions' effects should succeed, is core to ACID
guarantees. Although, at least by the ACID acronym, atomicity is not an
``isolation'' property, atomicity properties also restrict the updates
visible to other transactions. Accordingly, here, we consider the
\textit{isolation} effects of atomicity, which we call \textbf{Read
  Atomic (RA)} isolation.  Under RA, once some of the effects of a
transaction $T_i$ are observed by another transaction $T_j$,
thereafter, all effects of $T_i$ are observed by $T_j$. That is, if a
transaction $T_j$ reads a version of an object that transaction $T_i$
wrote, then a later read by $T_j$ cannot return value whose later
version is installed by $T_i$. Together with item cut isolation, RA
prevents Read Skew anomalies (Berenson et al.'s A5A) and is useful in
several contexts such as maintaining foreign key constraints,
consistent global secondary indexing, and maintenance of derived
data. In the example below, under RA, because $T_2$ has read $T_1$'s
write to $y$, $T_2$ must observe $b=c=1$ (or later versions for each
key):
\begin{align*}
\small
T_1 &: w_x(1)~w_y(1)~w_z(1)
\\T_2 &: r_x(a)~r_y(1)~r_x(b)~r_z(c)~\\[-1.5em]
\end{align*}
$T_2$ can also observe $a=\bot$, $a=1$, or a later version of $x$. In
the hierarchy of existing isolation properties, we place RA below
Adya's \textit{PL-2L} (as it does not necessarily enforce transitive
read-write dependencies) but above Read Committed ($PL-2$). Notably,
RA requires disallows reading intermediate writes (Adya's $G1b$):
observing all effects of a transaction implicitly requires observing
the final (committed) effects of the transaction as well.
